---
title: "Case Study 1"
author: "Smith Agarwal, Nitin Vashisth, Vindhya Singh, Akshaya Ramakrishnan"
date: "December 20, 2017"
output:
  beamer_presentation: default
  slidy_presentation: default
---

# Contents

- Data Loading
- Data Description
- Data Analysis - I
- Data Analysis - II
- Data Analysis - III

# Data Loading

```{r , message= FALSE}
#Loading the required libraries
library(data.table)
library(ggplot2)
library(tidyr)
library(dplyr)
```

```{r}
#Reading the input files and storing it in data table
file_path = "~/Desktop/TUM Course/Data Analysis and Visualization in R/Case Study/Data_sets"
gene <- fread(file.path(file_path,"gene.txt"))
genotype <- fread(file.path(file_path,"genotype.txt"))
growth <- fread(file.path(file_path,"growth.txt"))
marker <- fread(file.path(file_path,"marker.txt"))
expression <- fread(file.path(file_path,"expression.txt"))
```

# Data Description

```{r}
glimpse(gene)
dim(genotype)
glimpse(genotype[,1:5])
glimpse(growth)
glimpse(marker)
dim(expression)
glimpse(expression[,1:5])
```

# Data Analysis - I

####Analysing the fitness of segregants by plotting the growth rate of all segregants for different environments

```{r, tidy=TRUE, tidy.opts=list(width.cutoff=60), message=FALSE}
growth_gather <- gather(growth,key="env",value="growth_rate",-1)

ggplot(growth_gather,aes(x=env,y=growth_rate)) + geom_boxplot() + 
geom_jitter(width = 0.4,alpha=0.2)+labs(x="Environment",y="Growth Rate")

ggplot(growth_gather,aes(growth_rate)) + geom_density(aes(fill=env,alpha=0.25))+ labs(x="Growth Rate",y="Density")
```

Conclusion

- The fitness of the segregants vary greatly based on the environments
- Segregants display a maximum growth rate in YPD environment
- Within each environment, the fitness varies depending on the genotype of the segregants
- In YPD_BPS the genotype effect is smallest and in YPMalt the genotype affects fitness greatly

# Data Analysis - II

####Analysing the segregant fitness across all the markers by plotting the growth rate of all segregants for different environments and dominant strain

```{r, tidy=TRUE, tidy.opts=list(width.cutoff=60), message=FALSE}
# Calculating the count of strains for each marker grouped by strain values
genotype_mrk <- gather(genotype,"marker","value",-1)
genotype_mrk <- as.data.table(genotype_mrk)
genotype_strain_count <- genotype_mrk[,.(count=.N),by=.(value,strain)]
genotype_strain_count <- genotype_strain_count[,c(2,3,1)]
genotype_strain_count <- spread(genotype_strain_count,key=value,value=count)

#Differentiate a segment based on dominant strain
genotype_strain_count$genotype <- ifelse(genotype_strain_count$`Lab strain`>genotype_strain_count$`Wild isolate`,"Lab strain","Wild isolate")
growth_merged <- merge(growth_gather,genotype_strain_count)

ggplot(growth_merged,aes(x=env,y=growth_rate,color=genotype)) +labs(x="Environment",y="Growth Rate")+ geom_boxplot()
```

Conclusion

- We call a segregant 'Wild' when it has more wild markers then 'Lab' and vice versa
- In general, we observed that each segregant has almost 50% of the markers denoting 'Lab' and 'Wild' with only slight variations
- Except for YPE environment, the segregants with more 'Wild' markers have slightly better growth rate globally.
- IN YPD_BPS the genotype effect is smallest and in YPMalt the genotype affects fitness greatly

# Data Analysis - III

####Testing growth in Maltose and genotype given in the case study slides

```{r, tidy=TRUE, tidy.opts=list(width.cutoff=60), message=FALSE}
marker <- read.delim(file.path(file_path,"marker.txt"))
growth <- read.delim(file.path(file_path,"growth.txt"))
genotype <- read.delim(file.path(file_path,"genotype.txt"))
mygeno <- genotype[, which(marker$chrom=="chr07" & marker$start== 1069229)]
names(mygeno) <- genotype$strain
plot(YPMalt ~ mygeno[strain], data=growth)

#Finding the index of the desired marker
which(marker$chrom=="chr07" & marker$start== 1069229)
#Find the marker id for the obtained index
marker[which(marker$chrom=="chr07" & marker$start== 1069229),]
#Column name of the marker id used
colnames(genotype)[which(marker$chrom=="chr07" & marker$start== 1069229)]
#Desired column name of the marker id
colnames(genotype)[which(marker$chrom=="chr07" & marker$start== 1069229)+1]

#Desired Boxplot graph
mygeno <- genotype[, which(marker$chrom=="chr07" & marker$start== 1069229)+1]
names(mygeno) <- genotype$strain
plot(YPMalt ~ mygeno[strain], data=growth)
```

Conclusion

- The discrepancy was due to missing the index value by 1 as there is an additional column of strain in the genotype dataset making the number of variables as 1001
- Even though it was a different marker but the variation in both the graphs is minimal